---
title: "R Notebook"
---


```{r library-imports, results='hide', message=FALSE, include=FALSE, echo=FALSE}

# Time Series Analysis
library(tseries)
library(forecast)
library(TSstudio)
library(tswge)
library(orcutt)
library(vars)

# Date Manipulation
library(lubridate)

# Metrics
library(MLmetrics)

# Correlation Matrix
library(ggcorrplot)

# Misc
library(tidyverse)
library(kableExtra)
library(knitr)
library(Hmisc)

# Neural Networks
library(nnfor)

```


```{r load-data}

train <- read.csv("../../../data/train.csv")

test <- read.csv("../../../data/test.csv")

# copy the test and train data sets for EDA

train.c1 <- train
test.c1 <- test

train.c1$datetime = ymd_hms(train.c1$datetime)
test.c1$datetime = ymd_hms(test.c1$datetime)

train.c1 <- train.c1 %>%
  mutate(year = as.factor(format(datetime, format = "%Y")),
         # month = as.factor(format(datetime, format = "%m")),
         month = month(train.c1$datetime, label = TRUE, abbr = FALSE),
         day = as.factor(format(datetime, format = "%d")),
         hour = as.factor(format(datetime, format = "%H")),
         season = factor(season, labels = c("Spring", "Summer", "Fall", "Winter")),
         holiday = factor(holiday, labels = c("No", "Yes")),
         workingday = factor(workingday, labels = c("No", "Yes")),
         weather = factor(weather, labels = c("Great", "Good", "Average", "Poor")))


test.c1 <- test.c1 %>%
  mutate(year = as.factor(format(datetime, format = "%Y")),
         # month = as.factor(format(datetime, format = "%m")),
         month = month(test.c1$datetime, label = TRUE, abbr = FALSE),
         day = as.factor(format(datetime, format = "%d")),
         hour = as.factor(format(datetime, format = "%H")),
         season = factor(season, labels = c("Spring", "Summer", "Fall", "Winter")),
         holiday = factor(holiday, labels = c("No", "Yes")),
         workingday = factor(workingday, labels = c("No", "Yes")),
         weather = factor(weather, labels = c("Great", "Good", "Average", "Poor")))


```




  Column Name        |Type                  |Description
  -------------------|--------------------- |------------------------------------------------
  1.  datetime       |    Character         | YYYY-MM-DD HH24 (example:  2011-01-01 04:00:00)
  2.  season         |    Integer 		      | (1-4)
  3.  holiday        |    Integer 		      | (0 or 1)
  4.  workingday     |    Integer 		      | (0 or 1)
  5.  weather        |    Integer 		      | (1-4)
  6.  temp           |    Float 			      | temparture in Celcius
  7.  atemp          |    Float 			      | "feels like" temperature in Celsius
  8.  humidity       |    Integer           | relative humidity
  9.  windspeed      |    Float             | wind speed
  10. __casual__     |    Integer 		      | count of casual users 
  11. __registered__ |    Integer 		      | count of registered users 
  12. __count__      |    Integer 		      | count of total users (*primary response variable*)
  



```{r}
train.c1 %>%
  # mutate(price.mean = median(price)) +
  ggplot(aes(count)) +
  geom_histogram(bins = 150, alpha = 0.65) +
  ggtitle("Histogram of Hourly Bike Count") +
  labs(x = "Hourly Bike Count", y = "Totals per Bin") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

```{r}

summary(train.c1$count)

```



```{r}

train.c1 %>%
  remove_missing(na.rm = TRUE) %>%
  group_by(month) %>%
  ggplot(aes(x = month, y = count, fill=month)) +
  geom_bar(position = "dodge", stat = "summary", fun = "mean") +
  ggtitle("Bar Plot of Average Hourly Bike Rentals by Month") +
  labs(x = "Month", y = "Average") + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

```{r}
train.c1 %>%
  ggplot(aes(x=month, y=count, fill=month)) + 
  geom_boxplot() + 
  ggtitle("Box Plot of Bike Rentals by Month") +
  labs(x = "Month", y = "Average") + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```



```{r}

train.c1 %>%
  remove_missing(na.rm = TRUE) %>%
  group_by(season) %>%
  ggplot(aes(x = season, y = count, fill=season)) +
  geom_bar(position = "dodge", stat = "summary", fun = "mean") +
  ggtitle("Bar Plot of Average Hourly Bike Rentals by Season") +
  labs(x = "Season", y = "Average") + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

```{r}
train.c1 %>%
  ggplot(aes(x=season, y=count, fill=season)) + 
  geom_boxplot() + 
  ggtitle("Box Plot of Bike Rentals by Season") +
  labs(x = "Season", y = "Average") + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

```{r}

train.c1 %>%
  remove_missing(na.rm = TRUE) %>%
  group_by(hour) %>%
  ggplot(aes(x = hour, y = count, fill=hour)) +
  geom_bar(position = "dodge", stat = "summary", fun = "mean") +
  ggtitle("Bar Plot of Average Bike Rentals by Hour") +
  labs(x = "Hour", y = "Average") + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```
```{r}
train.c1 %>%
  ggplot(aes(x=hour, y=count, fill=hour)) + 
  geom_boxplot() + 
  ggtitle("Box Plot of Bike Rentals by Hour") +
  labs(x = "Hour", y = "Average") + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```



```{r}

train.c1 %>%
  remove_missing(na.rm = TRUE) %>%
  group_by(weather) %>%
  ggplot(aes(x = weather, y = count, fill=weather)) +
  geom_bar(position = "dodge", stat = "summary", fun = "mean") +
  ggtitle("Bar Plot of Average Hourly Bike Rentals by Weather") +
  labs(x = "Weather", y = "Average") + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

```{r}

train.c1 %>%
  ggplot(aes(x=weather, y=count, fill=weather)) + 
  geom_boxplot() + 
  ggtitle("Box Plot of Bike Rentals by Weather") +
  labs(x = "Weather", y = "Average") + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

```{r}

train.c1 %>%
  remove_missing(na.rm = TRUE) %>%
  group_by(workingday) %>%
  ggplot(aes(x = workingday, y = count, fill=workingday)) +
  geom_bar(position = "dodge", stat = "summary", fun = "mean") +
  ggtitle("Bar Plot of Average Hourly Bike Rentals by Working Day") +
  labs(x = "Working Day", y = "Average") + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```


```{r}

train.c1 %>%
  ggplot(aes(x=workingday, y=count, fill=workingday)) + 
  geom_boxplot() + 
  ggtitle("Box Plot of Bike Rentals by Working Day") +
  labs(x = "Working Day", y = "Average") + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

```{r}

train.c1 %>%
  remove_missing(na.rm = TRUE) %>%
  group_by(holiday) %>%
  ggplot(aes(x = holiday, y = count, fill=holiday)) +
  geom_bar(position = "dodge", stat = "summary", fun = "mean") +
  ggtitle("Bar Plot of Average Hourly Bike Rentals by Holiday") +
  labs(x = "Holiday", y = "Average") + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

```{r}

train.c1 %>%
  ggplot(aes(x=holiday, y=count, fill=holiday)) + 
  geom_boxplot() + 
  ggtitle("Box Plot of Bike Rentals by Holiday") +
  labs(x = "Holiday", y = "Average") + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

```{r}

train.c1.numeric <- train.c1 %>%
  select_if(is.numeric)

corr <- round(cor(train.c1.numeric), 1)
  
ggcorrplot(corr, method = "circle")


```

```{r}
train.c1 %>%
  ggplot(aes(x = temp, y = count)) + 
  geom_point(alpha = 0.3) + 
  geom_smooth() + 
  ggtitle("Line Chart of Counts by Tempurature") +
  labs(x = "Tempurature", y = "Count")
  
```

```{r}
train.c1 %>%
  ggplot(aes(x = atemp, y = count)) + 
  geom_point(alpha = 0.3) + 
  geom_smooth() + 
  ggtitle("Line Chart of Counts by 'Feels Like' Tempurature") +
  labs(x = "'Feels Like' Tempurature", y = "Count")
  
```

```{r}

train.c1 %>%
  ggplot(aes(x = temp, y = count)) + 
  geom_point(alpha = 0.3) + 
  geom_smooth() + 
  ggtitle("Line Chart of Counts by Wind Speed") +
  labs(x = "Wind Speed", y = "Count")
  
```

```{r}

train.c1 %>%
  ggplot(aes(x = humidity, y = count)) + 
  geom_point(alpha = 0.3) + 
  geom_smooth() + 
  ggtitle("Line Chart of Counts by Humidity") +
  labs(x = "Humidity", y = "Count")
  
```


```{r}

train.2012.june <-  train.c1 %>%
  filter(train.c1$year == 2012 & train.c1$month == "June")

```



```{r}

ggplot(data=train.2012.june, aes(x = datetime, y = count)) +
      geom_line() + 
      ggtitle("Hourly Rental Trends for June 1st - June 19th, 2012") +
      labs(x = "Day", y = "Count") + 
      scale_x_datetime(date_breaks = "1 day", date_labels = "%d")

```


```{r}

plotts.sample.wge(train.2012.june$count)

```

```{r}
Acf(train.2012.june$count, lag.max = 50)
```

```{r}

plotts.parzen.wge(train.2012.june$count)

```


```{r}

count.d1 = artrans.wge(train.2012.june$count, phi.tr = 1)

plotts.sample.wge(count.d1)


```

```{r}

# aic5.wge(train.2012.june$count, p=0:2, q=0:2, type = "aic") # BIC picks ARMA (15,0)
# aic5.wge(train.2012.june$count, p=24:26, q=0:2, type = "aic") # BIC picks ARMA (25,1)
# aic5.wge(train.2012.june$count, p=26:30, q=0:2, type = "bic") # BIC picks ARMA (26,0)

# aic5.wge(train.2012.june$count, p=24:26, q=0:5, type = "bic") # BIC picks ARMA (26,0)
# aic5.wge(train.2012.june$count, p=24:26, q=0:2, type = "aic") # BIC picks ARMA (26,4)

# aic5.wge(train.2012.june$count, p=0:15, q=0:2, type = "aic") # AIC picks ARMA (15,1)
# aic5.wge(train.2012.june$count, p=15:25, q=0:2, type = "aic") # AIC picks ARMA (24,2)
# aic5.wge(train.2012.june$count, p=26:30, q=0:2, type = "aic") # AIC picks ARMA (30, 1)

```


```{r}

# factor.wge(c=(rep(0,7), 1))

```



```{r}

est.2012.june = est.arma.wge(train.2012.june$count, p=25, q=1)

est.2012.june$phi
est.2012.june$theta
est.2012.june$avar

mean(train.2012.june$count)

```


$$\begin{equation}
   (1-0.80B+0.15B^2+0.04B^3-0.05B^4+0.02B^5+0.09B^6-0.09B^7+0.30B^8-0.34B^9+0.10B^{10}+0.12B^{11}-0.06B^{12}+0.03B^{13} + 0.05B^{14}-0.09B^{15}+0.20B^{16}-0.16B^{17}+0.05B^{18}+0.10B^{19}-0.06B^{20}-0.003B^{21}+0.06B^{22}-0.13B^{23}-0.45B^{24}+0.31B^{25})(X_t - 287.18) = (1+0.30B)a_t; \sigma^2 = 5623.31
\end{equation}$$


```{r}
plotts.wge(est.2012.june$res)
```


```{r}

# 7 day forecast

df <- train.2012.june

no_ahead = 24 * 7

len_of_obs = length(df$datetime)

forecast.2012.june = fore.aruma.wge(df$count, phi = est.2012.june$phi, theta = est.2012.june$theta, n.ahead = no_ahead, lastn = TRUE, limits = FALSE, plot = FALSE)

# plot(df$count[1:len_of_obs], type = "l", xlab = "Time", ylab = "Count", main = "7 Day Forecast (ARMA)")
# lines(seq(289,456), forecast.2012.june$f, col = "red")

ase.arma.forecasts.seven.days = mean((df$count[(len_of_obs - no_ahead + 1): len_of_obs] - forecast.2012.june$f)^2)
ase.arma.forecasts.seven.days

rmse.arma.forecasts.seven.days = RMSE(y_true = df$count[(len_of_obs - no_ahead + 1): len_of_obs], y_pred = forecast.2012.june$f)
rmse.arma.forecasts.seven.days


```

```{r}


forecasts.arma.25.1.seven.days = round(forecast.2012.june$f)

# set point forecast to 0 when less than 0
forecasts.arma.25.1.seven.days[forecasts.arma.25.1.seven.days < 0] <- 0

# predictions = round(forecast.2012.june$f)
# actuals = train.2012.june$count[(len_of_obs - no_ahead + 1): len_of_obs]
# 
# preds.v.actuals.2012.june <- data.frame(predictions, actuals)    # Apply data.frame function
# 
# 
# preds.v.actuals.2012.june <- preds.v.actuals.2012.june %>%
#   mutate(diff = (round(predictions) - actuals)) %>%
#   mutate(abs.diff = abs(round(predictions) - actuals))
# 
# preds.v.actuals.2012.june

```

```{r}

df2 <- data.frame(real = df$count,
          pred = c(rep(NA,length(df$datetime)-length(data.frame(forecasts.arma.25.1.seven.days)[,1])),data.frame(forecasts.arma.25.1.seven.days)[,1]),
          time = df$datetime)

plot.arma.seven.days <- ggplot(df2, aes(time, real, group = 1 )) +
  geom_line() + 
  geom_line(aes(time, pred, group = 1), color = "red")+
  ggtitle("7 Day Forecast (ARMA)") +
  labs(x = "Day", y = "Count") + 
  scale_x_datetime(date_breaks = "1 day", date_labels = "%d")

plot.arma.seven.days

```

```{r}

# 1 day forecast

df <- train.2012.june

no_ahead = 24

len_of_obs = length(df$datetime)

forecast.2012.june = fore.aruma.wge(df$count, phi = est.2012.june$phi, theta = est.2012.june$theta, n.ahead = no_ahead, lastn = TRUE, limits = FALSE, plot = FALSE)

# plot(df$count[1:len_of_obs], type = "l", xlab = "Time", ylab = "Count", main = "1 Day Forecast (ARMA)")
# lines(seq(433,456), forecast.2012.june$f, col = "red")

ase.arma.forecasts.one.day = mean((df$count[(len_of_obs - no_ahead + 1): len_of_obs] - forecast.2012.june$f)^2)
ase.arma.forecasts.one.day

rmse.arma.forecasts.one.day = RMSE(y_true = train.2012.june$count[(len_of_obs - no_ahead + 1): len_of_obs], y_pred = forecast.2012.june$f)
rmse.arma.forecasts.one.day


```

```{r}

forecasts.arma.25.1.one.day = round(forecast.2012.june$f)

forecasts.arma.25.1.one.day[forecasts.arma.25.1.one.day < 0] <- 0

# predictions = round(forecast.2012.june$f)
# actuals = train.2012.june$count[(len_of_obs - no_ahead + 1): len_of_obs]
# 
# preds.v.actuals.2012.june <- data.frame(predictions, actuals)    # Apply data.frame function
# 
# 
# preds.v.actuals.2012.june <- preds.v.actuals.2012.june %>%
#   mutate(diff = (round(predictions) - actuals)) %>%
#   mutate(abs.diff = abs(round(predictions) - actuals))
# 
# preds.v.actuals.2012.june

```

```{r}

df2 <- data.frame(real = df$count,
          pred = c(rep(NA,length(df$count)-length(data.frame(forecasts.arma.25.1.one.day)[,1])),data.frame(forecasts.arma.25.1.one.day)[,1]),
          time = df$datetime)

plot.arma.one.day <- ggplot(df2, aes(time, real, group = 1 )) +
  geom_line() + 
  geom_line(aes(time, pred, group = 1), color = "red")+
  ggtitle("1 Day Forecast (ARMA)") +
  labs(x = "Day", y = "Count") + 
  scale_x_datetime(date_breaks = "1 day", date_labels = "%d")


plot.arma.one.day

```





```{r}

# https://stackoverflow.com/questions/63359877/correlation-matrix-add-additional-column-of-adjusted-p-value

flattenCorrMatrix <- function(cormat, pmat) {
               ut <- upper.tri(cormat)
               data.frame(
               row = rownames(cormat)[row(cormat)[ut]],
               column = rownames(cormat)[col(cormat)[ut]],
               cor  =(cormat)[ut],
               p = pmat[ut]
               )
               }
```


```{r}

# library(Hmisc)

train.c1.dropna <- train.c1.numeric %>% drop_na() # %>% dplyr::select(-one_of(c('count')))
res2 <- rcorr(as.matrix(train.c1.dropna))
flattenedCor <- flattenCorrMatrix(res2$r, res2$P) %>% arrange(desc(cor))
flattenedCor <- flattenedCor %>% filter(abs(cor) > 0.1) %>% dplyr::select(-one_of(c('p')))
names(flattenedCor) <- c('Feature 1','Feature 2','Correlation Coefficient')
kable(flattenedCor)

```


```{r}

# series is the array of the series

# horizon is how far you want to predict into the future

# d is the order of the differencing: (1-B^)^d

# s is the order of the seasonality: (1-B^s)

# phis = order of the stationary AR term

# thetas = order of the invertible MA term

# It simply takes the given horizon and the model in the form of s,d,phis and
# thetas and figures out how many windows it can create in the data (series) and then calculates the ASE for each window.

#The output is the average off all the ASEs from each individual window.

roll.win.ase.wge = function(series, horizon = 1, s = 0, d = 0, phis = 0, thetas = 0) {
  
  trainingSize = length(phis) + length(thetas) + s + d + 1
  
  numwindows = length(series)-(trainingSize + horizon) + 1
  
  ASEHolder = numeric(numwindows)
  
  print(paste("Please Hold For a Moment, TSWGE is processing the Rolling Window ASE with", numwindows, "windows."))
  
  for( i in 1:numwindows) {
    invisible(capture.output(forecasts <- fore.aruma.wge(series[i:(i+(trainingSize-1))],phi = phis, theta = thetas, s = s, d = d,n.ahead = horizon)))
    ASE = mean((series[(trainingSize+i):(trainingSize+ i + (horizon) - 1)] - forecasts$f)^2)
    ASEHolder[i] = ASE
  }
  
  ASEHolder
  hist(ASEHolder, main = "ASEs for Individual Windows")
  WindowedASE = mean(ASEHolder)
  print("The Summary Statistics for the Rolling Window ASE Are:")
  print(summary(ASEHolder))
  print(paste("The Rolling Window ASE is: ",round(WindowedASE,3)))
  
  return(list(rwASE = WindowedASE, numwindows = numwindows, horizon = horizon, s = s, d = d, phis = phis, thetas = thetas))
}


```


```{r}

# roll.win.ase.wge(train.2012.june$count, horizon = 24, est.2012.june$phi, est.2012.june$theta, s = 0, d = 0)

```



# Vector Auto-Regressive (VAR) Model

```{r}
ccf(train.2012.june$count, train.2012.june$temp)
```

```{r}

ccf(train.2012.june$count, train.2012.june$weather)

```

```{r}

ccf(train.2012.june$count, train.2012.june$humidity)

```

```{r}

ccf(train.2012.june$count, train.2012.june$hour)

```

```{r}

# train.2012.june$count <- log(train.2012.june$count)

# train.2012.june$temp <- scale(train.2012.june$temp, center = FALSE, scale = TRUE) 
# train.2012.june$humidity <- scale(train.2012.june$humidity, center = FALSE, scale = TRUE) 
# train.2012.june$windspeed <- scale(train.2012.june$windspeed, center = FALSE, scale = TRUE) 

X = cbind(train.2012.june$count, train.2012.june$temp, train.2012.june$humidity, train.2012.june$weather, train.2012.june$hour)
# X = cbind(train.c1$count, train.c1$temp, train.c1$humidity, train.c1$weather, train.c1$hour)

VARselect(X, lag.max = 24, type = "const", season = NULL, exogen = NULL)

```


```{r}

lsfit = VAR(X, p = 22, type = "const", ic = c("AIC"))
# summary(lsfit)

```

## 7 Day Forecast

```{r}

no_ahead = 24*7

forecasts.var.seven.days = predict(lsfit, n.ahead = no_ahead)

```


```{r}

forecasts.1.var.seven.days = forecasts.var.seven.days$fcst$y1[,1]

# set point forecast to 0 when less than 0
forecasts.1.var.seven.days[forecasts.1.var.seven.days < 0] <- 0

```

```{r}

# no_ahead = 24*7

df <- train.2012.june

len_of_obs = length(df$datetime)

train_stop = len_of_obs - no_ahead
test_start = train_stop + 1

plot(df$count[1:len_of_obs], type = "l", xlab = "Time", ylab = "Count", main = "7 Day Forecast (VAR)")
lines(seq(test_start, len_of_obs), forecasts.1.var.seven.days, col = "red")

ase.var.forecasts.seven.days = mean((df$count[test_start:len_of_obs] - forecasts.1.var.seven.days)^2)
ase.var.forecasts.seven.days

rmse.var.2012.seven.days = RMSE(y_true = df$count[(len_of_obs - no_ahead + 1): len_of_obs], y_pred = forecasts.1.var.seven.days)
rmse.var.2012.seven.days

```


```{r}

df2 <- data.frame(real = df$count,
          pred = c(rep(NA,length(df$count)-length(data.frame(forecasts.1.var.seven.days)[,1])),data.frame(forecasts.1.var.seven.days)[,1]),
          time = df$datetime)

plot.arma.one.day <- ggplot(df2, aes(time, real, group = 1 )) +
  geom_line() + 
  geom_line(aes(time, pred, group = 1), color = "red")+
  ggtitle("7 Day Forecast (VAR)") +
  labs(x = "Day", y = "Count") + 
  scale_x_datetime(date_breaks = "1 day", date_labels = "%d")


plot.arma.one.day

```



## 1 Day Forecast

```{r}

no_ahead = 24

forecasts.var.one.day = predict(lsfit, n.ahead = no_ahead)

```


```{r}

forecasts.1.var.one.day = forecasts.var.one.day$fcst$y1[,1]

# set point forecast to 0 when less than 0
forecasts.1.var.one.day[forecasts.1.var.one.day < 0] <- 0

```

```{r}

# no_ahead = 24

df <- train.2012.june

len_of_obs = length(df$datetime)

train_stop = len_of_obs - no_ahead
test_start = train_stop + 1

# plot(df$count[1:len_of_obs], type = "l", xlab = "Time", ylab = "Count", main = "1 Day Forecast (VAR)")
# lines(seq(test_start, len_of_obs), forecasts.1.var.one.day, col = "red")

ase.var.forecasts.one.day = mean((df$count[test_start:len_of_obs] - forecasts.1.var.one.day)^2)
ase.var.forecasts.one.day

rmse.var.forecasts.one.day = RMSE(y_true = df$count[(len_of_obs - no_ahead + 1): len_of_obs], y_pred = forecasts.1.var.one.day)
rmse.var.forecasts.one.day

```

```{r}

df2 <- data.frame(real = df$count,
          pred = c(rep(NA,length(df$count)-length(data.frame(forecasts.1.var.one.day)[,1])),data.frame(forecasts.1.var.one.day)[,1]),
          time = df$datetime)

plot.arma.one.day <- ggplot(df2, aes(time, real, group = 1 )) +
  geom_line() + 
  geom_line(aes(time, pred, group = 1), color = "red")+
  ggtitle("1 Day Forecast (VAR)") +
  labs(x = "Day", y = "Count") + 
  scale_x_datetime(date_breaks = "1 day", date_labels = "%d")


plot.arma.one.day

```


# Neural Network Model

## 7 Day Forecast

```{r}

df <- train.2012.june

no_ahead = 24 * 7

len_of_obs = length(df$datetime)

train_stop = len_of_obs - no_ahead
test_start = train_stop + 1

# df$temp <- scale(df$temp, center = FALSE, scale = TRUE)
# df$humidity <- scale(df$humidity, center = FALSE, scale = TRUE)
# df$windspeed <- scale(df$windspeed, center = FALSE, scale = TRUE)

ts.count = ts(df$count[1:train_stop])
ts.count.x = data.frame(temp = ts(df$temp), humidity = ts(df$humidity),
                        weather = ts(df$weather), hour = ts(df$hour))

set.seed(42)

fit.mlp.seven.days = mlp(ts.count, hd = 5, xreg = ts.count.x, lags = no_ahead, xreg.lags = c(1,1,1,24), allow.det.season = FALSE)
plot(fit.mlp.seven.days)

forecasts.mlp.seven.days = forecast(fit.mlp.seven.days, h = no_ahead, xreg = ts.count.x)

# plot(df$count[1:len_of_obs], type = "l", xlab = "Time", ylab = "Count", main = "7 Day Forecast (MLP)")
# lines(seq(test_start, len_of_obs), forecasts.mlp.seven.days$mean, col = "red")

ase.mlp.forecasts.seven.days = mean((df$count[test_start:len_of_obs] - forecasts.mlp.seven.days$mean)^2)
ase.mlp.forecasts.seven.days

rmse.mlp.forecasts.seven.days = sqrt(mean((df$count[test_start:len_of_obs] - forecasts.mlp.seven.days$mean)^2))
rmse.mlp.forecasts.seven.days


```

```{r}

df2 <- data.frame(real = df$count,
          pred = c(rep(NA,length(df$count)-length(data.frame(forecasts.mlp.seven.days)[,1])),data.frame(forecasts.mlp.seven.days)[,1]),
          time = df$datetime)

plot.arma.one.day <- ggplot(df2, aes(time, real, group = 1 )) +
  geom_line() + 
  geom_line(aes(time, pred, group = 1), color = "red")+
  ggtitle("7 Day Forecast (MLP)") +
  labs(x = "Day", y = "Count") + 
  scale_x_datetime(date_breaks = "1 day", date_labels = "%d")


plot.arma.one.day

```


## 1 Day Forecast

```{r}

df <- train.2012.june

no_ahead = 24

len_of_obs = length(df$datetime)

train_stop = len_of_obs - no_ahead
test_start = train_stop + 1

# df$temp <- scale(df$temp, center = FALSE, scale = TRUE)
# df$humidity <- scale(df$humidity, center = FALSE, scale = TRUE)
# df$windspeed <- scale(df$windspeed, center = FALSE, scale = TRUE)

ts.count = ts(df$count[1:train_stop])
ts.count.x = data.frame(temp = ts(df$temp), humidity = ts(df$humidity), 
                        weather = ts(df$weather), hour = ts(df$hour))
set.seed(42)

# xreg.lags = c(1,1,1,24),

fit.mlp.one.day = mlp(ts.count, hd = 5, xreg = ts.count.x, lags = no_ahead, allow.det.season = FALSE)
plot(fit.mlp.one.day)

forecasts.mlp.one.day = forecast(fit.mlp.one.day, h = no_ahead, xreg = ts.count.x)

# plot(df$count[1:len_of_obs], type = "l", xlab = "Time", ylab = "Count", main = "1 Day Forecast (MLP)")
# lines(seq(test_start, len_of_obs), forecasts.mlp.one.day$mean, col = "red")

ase.mlp.forecasts.one.day = mean((df$count[test_start:len_of_obs] - forecasts.mlp.one.day$mean)^2)
ase.mlp.forecasts.one.day

rmse.mlp.forecasts.one.day = sqrt(mean((df$count[test_start:len_of_obs] - forecasts.mlp.one.day$mean)^2))
rmse.mlp.forecasts.one.day

```

```{r}

df2 <- data.frame(real = df$count,
          pred = c(rep(NA,length(df$count)-length(data.frame(forecasts.mlp.one.day)[,1])),data.frame(forecasts.mlp.one.day)[,1]),
          time = df$datetime)

plot.arma.one.day <- ggplot(df2, aes(time, real, group = 1 )) +
  geom_line() + 
  geom_line(aes(time, pred, group = 1), color = "red")+
  ggtitle("1 Day Forecast (MLP)") +
  labs(x = "Day", y = "Count") + 
  scale_x_datetime(date_breaks = "1 day", date_labels = "%d")


plot.arma.one.day

```


## Ensemble 

### 7 Days

```{r}

forecasts.1.var.seven.days 
forecasts.mlp.seven.days

df.ensemble.seven.days <- data.frame(train.2012.june$count[289:456], forecasts.arma.25.1.seven.days, list(forecasts.1.var.seven.days), forecasts.mlp.seven.days)

colnames(df.ensemble.seven.days) <- c("Actual", "Predicted (ARMA)", "Predicted (VAR)", "Predicted (MLP)")

df.ensemble.seven.days <- df.ensemble.seven.days %>%
  mutate("Predicted (VAR)" = round(df.ensemble.seven.days$`Predicted (VAR)`)) %>%
  mutate("Predicted (MLP)" = round(df.ensemble.seven.days$`Predicted (MLP)`)) %>%
  mutate("Predicted (Ensemble)" = round((df.ensemble.seven.days$`Predicted (MLP)` + df.ensemble.seven.days$`Predicted (VAR)`)/2))

df.ensemble.seven.days

```

```{r}

mean((df.ensemble.seven.days$`Actual` - df.ensemble.seven.days$`Predicted (ARMA)`)^2)

mean((df.ensemble.seven.days$`Actual` - df.ensemble.seven.days$`Predicted (VAR)`)^2)

mean((df.ensemble.seven.days$`Actual` - df.ensemble.seven.days$`Predicted (MLP)`)^2)

mean((df.ensemble.seven.days$`Actual` - df.ensemble.seven.days$`Predicted (Ensemble)`)^2)

```

```{r}

sqrt(mean((df.ensemble.seven.days$`Actual` - df.ensemble.seven.days$`Predicted (ARMA)`)^2))

sqrt(mean((df.ensemble.seven.days$`Actual` - df.ensemble.seven.days$`Predicted (VAR)`)^2))

sqrt(mean((df.ensemble.seven.days$`Actual` - df.ensemble.seven.days$`Predicted (MLP)`)^2))

sqrt(mean((df.ensemble.seven.days$`Actual` - df.ensemble.seven.days$`Predicted (Ensemble)`)^2))

```

# 1 Day

```{r}

forecasts.1.var.one.day # = data.frame(forecasts.1.var.seven.days)
forecasts.mlp.one.day

df.ensemble.one.day <- data.frame(train.2012.june$count[433:456], forecasts.arma.25.1.one.day, list(forecasts.1.var.one.day), forecasts.mlp.one.day)

colnames(df.ensemble.one.day) <- c("Actual", "Predicted (ARMA)", "Predicted (VAR)", "Predicted (MLP)")

df.ensemble.one.day <- df.ensemble.one.day %>%
  mutate("Predicted (VAR)" = round(df.ensemble.one.day$`Predicted (VAR)`)) %>%
  mutate("Predicted (MLP)" = round(df.ensemble.one.day$`Predicted (MLP)`)) %>%
  mutate("Predicted (Ensemble)" = round((df.ensemble.one.day$`Predicted (MLP)` + df.ensemble.one.day$`Predicted (VAR)`)/2))

df.ensemble.one.day

```
### ASE 

```{r}


mean((df.ensemble.one.day$`Actual` - df.ensemble.one.day$`Predicted (ARMA)`)^2)

mean((df.ensemble.one.day$`Actual` - df.ensemble.one.day$`Predicted (VAR)`)^2)

mean((df.ensemble.one.day$`Actual` - df.ensemble.one.day$`Predicted (MLP)`)^2)

mean((df.ensemble.one.day$`Actual` - df.ensemble.one.day$`Predicted (Ensemble)`)^2)

```


## RMSE

```{r}


sqrt(mean((df.ensemble.one.day$`Actual` - df.ensemble.one.day$`Predicted (ARMA)`)^2))

sqrt(mean((df.ensemble.one.day$`Actual` - df.ensemble.one.day$`Predicted (VAR)`)^2))

sqrt(mean((df.ensemble.one.day$`Actual` - df.ensemble.one.day$`Predicted (MLP)`)^2))

sqrt(mean((df.ensemble.one.day$`Actual` - df.ensemble.one.day$`Predicted (Ensemble)`)^2))

```












